---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "images/"
)
```

# readtext: Import and handling for plain and formatted text files

[![CRAN Version](http://www.r-pkg.org/badges/version/readtext)](https://CRAN.R-project.org/package=readtext)
![Downloads](http://cranlogs.r-pkg.org/badges/readtext)
[![Travis-CI Build Status](https://travis-ci.org/kbenoit/readtext.svg?branch=master)](https://travis-ci.org/kbenoit/readtext)
[![Build status](https://ci.appveyor.com/api/projects/status/x6dtvh2m7mj3b026/branch/master?svg=true)](https://ci.appveyor.com/project/kbenoit/readtext)
[![codecov.io](https://codecov.io/github/kbenoit/readtext/coverage.svg?branch=master)][1]


[1]: https://codecov.io/gh/kbenoit/readtext/branch/master

An R package for reading text files in all their various formats, by Ken Benoit, Adam Obeng, Paul Nulty, and Stefan Müller.

## Introduction

**readtext** is a one-function package that does exactly what it says on the tin: It reads files containing text, along with any associated document-level metadata, which we call "docvars", for document variables.  Plain text files do not have docvars, but other forms such as .csv, .tab, .xml, and .json files usually do.  

**readtext** accepts filemasks, so that you can specify a pattern to load multiple texts, and these texts can even be of multiple types.  **readtext** is smart enough to process them correctly, returning a data.frame with a primary field "text" containing a character vector of the texts, and additional columns of the data.frame as found in the document variables from the source files.

As encoding can also be a challenging issue for those reading in texts, we include functions for diagnosing encodings on a file-by-file basis, and allow you to specify vectorized input encodings to read in file types with individually set (and different) encodings.  (All encoding functions are handled by the **stringi** package.)

## How to Install

1.  From GitHub

    ```{r, eval = FALSE}
    # devtools packaged required to install readtext from Github 
    devtools::install_github("kbenoit/readtext") 
    ```

2.  From CRAN

    ```{r, eval = FALSE}
    install.packages("readtext")
    ```


## Demonstration: Reading one or more text files

**readtext** supports plain text files (.txt), data in some form of JavaScript Object Notation (.json), comma-or tab-separated values (.csv, .tab, .tsv), XML documents (.xml), as well as PDF and Microsoft Word formatted files (.pdf, .doc, .docx). **readtext** also handles multiple files and file types using for instance a "glob" expression, files from a URL or an archive file (.zip, .tar, .tar.gz, .tar.bz).

Usually, you do not have to determine the format of the files explicitly - readtext takes this information from the file ending.

    ```{r, eval = FALSE}
    # get the data directory from readtext
    DATA_DIR <- system.file("extdata/", package = "readtext")
    
    # read in all files from a folder
    readtext(paste0(DATA_DIR, "/txt/UDHR/*"))
    
    # read in comma-separated values and specify text field
    readtext(paste0(DATA_DIR, "/csv/inaugCorpus.csv"), text_field = "texts")
    ```

## Inter-operability with **quanteda**

**readtext** was originally developed in early versions of the [**quanteda**](http:/github.com/kbenoit/quanteda) package for the quantitative analysis of textual data. Because quanteda's corpus constructor recognizes the data.frame format returned by `readtext()`, it can construct a corpus directly from a readtext object, preserving all docvars and other meta-data.

    ```{r, eval = FALSE}
    require(quanteda)
    
    # read in comma-separated values with readtext
    rt_csv <- readtext(paste0(DATA_DIR, "/csv/inaugCorpus.csv"), text_field = "texts")
    
    # create quanteda corpus
    corpus_csv <- corpus(rt_csv)
    summary(corpus_csv, 5)
    ```

The **readtext** [vignette](https://github.com/kbenoit/readtext/blob/master/inst/doc/readtext_vignette.Rmd) explains in more detail how to use the package. 


